---
title: "Quality Features Linear Regressions Analysis"
author: "NYC Data Science Academy"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message=FALSE)
```

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
```

1. Load: Load the Quality Features dataset into your workspace.
```{r}
quality = read.csv('./quality.csv')
quality = quality %>% select(-c(X))
head(quality)
```

2. Categorical variables -> Factors with baseline features set in level: 
```{r}
colnames(quality)
```
```{r}
quality$OverallQual=factor(quality$OverallQual, levels=c(5,6,7,8,4,9,3,10,2,1))
quality$BldgType=factor(quality$BldgType, levels=c('1Fam','TwnhsE','Twnhs','Duplex','2fmCon'))
quality$HouseStyle=factor(quality$HouseStyle, levels=c('1Story','2Story','1.5Fin','SLvl','SFoyer','2.5Unf','1.5Unf','2.5Fin'))
quality$ExterQual=factor(quality$ExterQual, levels=c('TA','Gd','Ex','Fa'))
quality$KitchenQual=factor(quality$KitchenQual, levels=c('TA','Gd','Ex','Fa','Po'))
quality$BsmtQual=factor(quality$BsmtQual, levels=c('TA','Gd','Ex','Fa','No Basement','Po'))
quality$BsmtExposure=factor(quality$BsmtExposure, levels=c('No','Av','Gd','Mn','No Basement','Missing'))
quality$Foundation=factor(quality$Foundation, levels=c('CBlock','PConc','BrkTil','Slab','Stone','Wood'))
quality$GarageType=factor(quality$GarageType, levels=c('Attchd','Detchd','BuiltIn','No Garage','Basment','2Types','CarPort'))
quality$GarageFinish=factor(quality$GarageFinish, levels=c('Unf','RFn','Fin','No Garage','Missing'))
quality$Exterior1st=factor(quality$Exterior1st, levels=c('VinylSd','HdBoard','MetalSd','Wd Sdng','Plywood','CemntBd','BrkFace','WdShing','Stucco','AsbShng','BrkComm','AsphShn','CBlock','PreCast','ImStucc'))
```

```{r}
class(quality$OverallQual) 
```
Take the log of the response y-variable:
```{r}
quality$SalePrice=log(quality$SalePrice)
head(quality)
```

3. Fit a Model: Fit a multiple linear regression predicting the price of a house based on the factorized and Year variables
For this model:
```{r}
colnames(quality)
```
```{r}
model.saturated = lm(SalePrice ~ OverallQual + YearBuilt + YearRemodAdd + BldgType + HouseStyle + ExterQual + KitchenQual + BsmtQual + BsmtExposure + Foundation + GarageType + GarageYrBlt + GarageFinish + Exterior1st, data = quality)
summary(model.saturated)
```
Choose: OverallQual, KitchenQual?, BsmtQual?, BsmtExposure?, GarageType 

3. Fit a Model: Fit a multiple linear regression predicting the price of a house based on the factorized and Year variables
For this model:
a. Write out the regression equation.
b. Interpret the meaning each of the 5 coefficients in context of the problem.
c. Are the coefficients significant? How can you tell?
d. Is the overall regression significant? How can you tell?
e. Find and interpret the RSE.
f. Find and interpret the adjusted coefficient of determination.
```{r}
# model.saturated = lm(Price ~ Food + Decor + Service + Location, data = restaurants)
# summary(model.saturated)

#The regression equation is: Price = -21.96 + 1.54*Food + 1.91*Decor - 0.003*Service
#- 2.07*West.
#
#When all ratings are 0 and the restaurant is on the East side, the
#average expected price of dinner is about $-21.96 (this does not make sense in
#context of this problem).
#
#As the food rating for a restaurant increases by one
#point, the price increases on average by about $1.54 WHILE HOLDING ALL OTHER
#VARIABLES CONSTANT.
#
#As the decor rating for a restaurant increases by one
#point, the price increases on average by about $1.91 WHILE HOLDING ALL OTHER
#VARIABLES CONSTANT.
#
#As the service rating for a restaurant increases by one
#point, the price decreases on average by about $0.003 WHILE HOLDING ALL OTHER
#VARIABLES CONSTANT.
#
#A restaurant on the west side has an average dinner price that is about $2.07
#cheaper than a meal on the east side WHILE HOLDING ALL OTHER VARIABLES CONSTANT.
#
#All coefficients are significant except for the Service coefficient.
#
#The overall F-statistic is significant so the overall regression is significant.
#
#The RSE is about 5.738, which is an estimate of the average deviation of the
#observations around the regression line.
#
#The adjusted R^2 is 0.6187, meaning that approximately 61.87% of the variation
#in dinner price is accounted for by the variables in our model.

```

4. Diagnostics: Investigate the assumptions of the model using the plot() function. Are there any violations?
```{r}
plot(model.saturated)

#No overt deviations from any of the assumptions.

```

5. Outliers: Investigate the influence plot for the model. Are there any restaurants about which we should be concerned?
```{r}
library(car)
influencePlot(model.saturated)

#There are a few points that surface with either high residuals or high hat-values,
#but there are none that have a severe negative influence on the regression surface.
```

6. Multicollinearity: Investigate the coefficient variance inflation factors; use these values to discuss multicollinearity.
```{r}
vif(model.saturated)

#The VIF of Service is about 3.6, which is decently high. As we saw in the
#scatterplot matrix above, we might be cautious about multicollinearity in our
#data because of correlations among our predictor variables.
```

7. A-V Plots: Create added variable plots for this model. What conclusions might you draw from these plots?
```{r}
avPlots(model.saturated)

#The food, decor, and location variables all seem to add some type of information
#to our model; however, the service variable doesn't seem to add any information
#when among the other predictors in this model.
```
SLR: 
```{r} 
#Year features 
model1.saturated = lm(SalePrice ~ YearBuilt + YearRemodAdd + GarageYrBlt, data = quality)
summary(model1.saturated)
plot(model1.saturated)
influencePlot(model1.saturated)
vif(model1.saturated)
avPlots(model1.saturated)
```

```{r}
# OverAllQual
model2.saturated = lm(SalePrice ~ OverallQual, data = quality)
summary(model2.saturated)
plot(model2.saturated)
influencePlot(mmodel2.saturatedodel2)
vif(model2.saturated)
avPlots(model2.saturated)
```
```{r}
# KitchenQual  
model3.saturated = lm(SalePrice ~ KitchenQual, data = quality)
summary(model3.saturated)
plot(model3.saturated)
influencePlot(mmodel3.saturatedodel2)
vif(model3.saturated)
avPlots(model3.saturated)
```


```{r}
# BsmtQual  
model4.saturated = lm(SalePrice ~ BsmtQual, data = quality)
summary(model4.saturated)
plot(model4.saturated)
influencePlot(mmodel4.saturatedodel2)
vif(model4.saturated)
avPlots(model4.saturated)
```

```{r}
#BsmtExposure
model5.saturated = lm(SalePrice ~ BsmtExposure, data = quality)
summary(model5.saturated)
plot(model5.saturated)
influencePlot(mmodel5.saturatedodel2)
vif(model5.saturated)
avPlots(model5.saturated)
```

```{r}
#GarageType
model6.saturated = lm(SalePrice ~ GarageType, data = quality)
summary(model6.saturated)
plot(model6.saturated)
influencePlot(mmodel6.saturatedodel2)
vif(model6.saturated)
avPlots(model6.saturated)
```

```{r}
#HouseStyle
model7.saturated = lm(SalePrice ~ HouseStyle, data = quality)
summary(model7.saturated)
plot(model7.saturated)
influencePlot(mmodel7.saturatedodel2)
vif(model7.saturated)
avPlots(model7.saturated)
```


```{r}
#BldgType
model8.saturated = lm(SalePrice ~ BldgType, data = quality)
summary(model8.saturated)
plot(model8.saturated)
influencePlot(mmodel8.saturatedodel2)
vif(model8.saturated)
avPlots(model8.saturated)
```


```{r}
#ExterQual
model9.saturated = lm(SalePrice ~ ExterQual, data = quality)
summary(model9.saturated)
```


```{r}
#Foundation
model10.saturated = lm(SalePrice ~ Foundation, data = quality)
summary(model10.saturated)
```


```{r}
#GarageFinish
model11.saturated = lm(SalePrice ~ GarageFinish, data = quality)
summary(model11.saturated)
```


```{r}
#Exterior1st
model12.saturated = lm(SalePrice ~ Exterior1st, data = quality)
summary(model12.saturated)
```


```{r}
#YearBuilt
model13.saturated = lm(SalePrice ~ YearBuilt, data = quality)
summary(model13.saturated)
```


```{r}
#YearRemodAdd
model14.saturated = lm(SalePrice ~ YearRemodAdd, data = quality)
summary(model14.saturated)
```


```{r}
#GarageYrBlt
model15.saturated = lm(SalePrice ~ GarageYrBlt, data = quality)
summary(model15.saturated)
```
Based on the MLR results, the top ones are: OverallQual, BsmtQual, ExternalQual, KitchenQual + in terms of Year feature, YearBuilt has higher R^2 for alone but YearModel has higher p-value and coef estimate per unit increase.  

```{r}
colnames(quality)
```

2. Compare: Using the metrics described in Section 1 above (e.g. R-squared, RSE, coefficient magnitudes/significance, AIC/BIC, etc.) compare your different models. Which do you think is best?
```{r}
### FINAL Model comparison ###
#AIC

AIC(model.saturated, model1, model2, model.service)

#BIC
BIC(model.saturated, model1, model2, model.service)

#model 2 has the least AIC, model 1 has the least BIC. All are relatively close,
#so you could make an argument for any one of them.
```
